### 2021-06-21 supp_file/supp_file.Rmd
as

---
title: "Supplemental File"
author: "Maxwel Coura Oliveira"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)
```

```{r message=FALSE, include=FALSE}
# Packages
options(scipen = 999)
library(tidyverse)
library(kableExtra)

```

```{r}
read_csv("bf.csv") %>% 
  kableExtra::kable(., booktabs = T, longtable = T, 
                    caption = "Bayes Factor interpretation",
                    linesep = "") %>%
  kable_styling(latex_options = c("striped", "repeat_header")) 
```

```{r echo=FALSE}
read_csv("inswath_bf.csv") %>% 
  mutate_if(is_character, ~str_replace(., "[)]", "")) %>% 
  mutate_if(is_character, ~str_replace(., "[(]", "")) %>% 
  kableExtra::kable(., booktabs = T, longtable = T, 
                    caption = "Treatment contrasts for inswath spray solution deposition. Bayes Factor is used to test contrast hypothesis",
                    linesep = "") %>%
  kable_styling(latex_options = c("striped", "repeat_header")) 
```

## AUC Bayes Factor contrasts

```{r echo=FALSE}
read_csv("auc_bf.csv") %>% 
  mutate_if(is.double, ~round(., 2)) %>% 
  mutate_if(is_character, ~str_replace(., "[)]", "")) %>% 
  mutate_if(is_character, ~str_replace(., "[(]", "")) %>% 
  kableExtra::kable(., booktabs = T, longtable = T, 
                    caption = "Treatment contrasts for Area Under the Curve. Bayes Factor is used to test contrast hypothesis",
                    linesep = "") %>%
  kable_styling(latex_options = c("striped", "repeat_header")) 
```