### 2021-06-21 analysis/droplet.Rmd

---
title: "droptlet"
author: "Maxwel Coura Oliveira"
date: "6/10/2021"
output: html_document
---

```{r}
library(tidyverse)
library(glmmTMB)
library(emmeans)
library(betareg)
library(lme4)
library(multcomp)
library(lmerTest)
library(car)
library(brms)
```

```{r}
readxl::read_excel("droplet_data.xlsx")
```

```{r}
readxl::read_excel("droplet_size_data.xlsx") %>% 
  janitor::clean_names() %>% 
  mutate_if(is_character, as_factor) %>% 
  group_by(solution, nozzle) %>% 
  mutate(nozzle = fct_recode(nozzle, 
                             "ULD12003" = "ULD11003")) %>% 
  mutate(x200 = x200/100) %>% 
  mutate(rep = row_number()) -> droplet
```

# DV10

```{r}
model1 <- glm(`dv10` ~ solution * nozzle,
                  family = Gamma(link = "inverse"),
                  data = droplet)
```

```{r}
Anova(model1)
```

```{r}
emmeans1 <- emmeans(model1, ~ nozzle * solution, type = "response")
```

```{r}
cld1 <- cld(emmeans1, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>% 
  as_tibble()
cld1
```

# DV50

```{r}
model2 <- glm(`dv50` ~ solution * nozzle,
                  family = Gamma(link = "inverse"),
                  data = droplet)
```

```{r}
Anova(model2)
```

```{r}
emmeans2 <- emmeans(model2, ~ nozzle * solution, type = "response")
```

```{r}
cld2 <- cld(emmeans2, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>% 
  as_tibble()
cld2
```

# DV90

```{r}
model3 <- glm(`dv90` ~ solution * nozzle,
                  family = Gamma(link = "inverse"),
                  data = droplet)
```

```{r}
Anova(model3)
```

```{r}
emmeans3 <- emmeans(model3, ~ nozzle * solution, type = "response")
```

```{r}
cld3 <- cld(emmeans3, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>% 
  as_tibble()
cld3
```

# RS

```{r}
model4 <- glm(rs ~ solution * nozzle,
                  family = Gamma(link = "inverse"),
                  data = droplet)
```

```{r}
Anova(model4)
```

```{r}
emmeans4 <- emmeans(model4, ~ nozzle * solution, type = "response")
```

```{r}
cld4 <- cld(emmeans4, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>% 
  as_tibble()
cld4
```

# x200

```{r}
model5 <- betareg(x200 ~ solution * nozzle, 
                  link = "logit",
                  data = droplet)
```

```{r}
Anova(model5)
```

```{r}
emmeans5 <- emmeans(model5, ~ nozzle * solution, type = "response")
```

```{r}
cld5 <- cld(emmeans5, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>% 
  as_tibble() 
cld5
```