### 2021-06-22 manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
  a4paper,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{Arial}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Spray particle drift},
  pdfauthor={Maxwel C. Oliveira, Guilherme Alves, Greg Kruger and Rodrigo Werle},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin = 2.54cm]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage[left]{lineno}
\linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Spray particle drift}
\author{Maxwel C. Oliveira, Guilherme Alves, Greg Kruger and Rodrigo
Werle}
\date{2021-06-22}

\begin{document}
\maketitle
\begin{abstract}
BACKGROUND:
\end{abstract}

\setstretch{2}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

Solution, sprayer and nozzle factors were grouped as a single fixed
effect (herein treatments) due to missing factor water in Missouri.
Resulting in a combination of 12 treatments.

\hypertarget{statistical-analyses}{%
\subsection{Statistical analyses}\label{statistical-analyses}}

The statistical analyses were conducted with R statistical software
version 4.1.0.\textsuperscript{1} Data analyses were performed with
Bayesian inference with ``brms'' package.\textsuperscript{2} Bayesian
inference uses Markov chain Monte Carlo algorithms for sampling a
probability distribution;\textsuperscript{2} and avoid singular fit from
frequentist linear models when using complex random effects.

\hypertarget{spray-solution-deposition-at-upwind-and-inswath}{%
\subsubsection{Spray solution deposition at upwind and
inswath}\label{spray-solution-deposition-at-upwind-and-inswath}}

Data was fitted to a mixed model using \emph{brm} function. Treatments
were the fixed effects and block nested within location random effects.
Model family was gaussian and prior distribution was set to student-t
with mean 0.5, standard deviation 3 and 11 degrees of freedom. The
posterior summaries (mean and highest posterior density) were estimated
with \emph{emmeans} function from ``emmeans'' package. Treatment means
were compared using Bayes Factor (BF).\textsuperscript{3,4} In short, if
BF \textgreater{} 1 there is evidence for H1 (difference between
treatments); whereas, if H0 \textless{} 1 there is evidence for H0 (no
difference between treatments). If BF = 1, there is no evidence. The
level of evidence (anecdotal, moderate, strong, very strong, and
extreme) varies as the BF value increase (evidence for H1) or decrease
(evidence for H0).

\hypertarget{spray-solution-deposition-at-downwind}{%
\subsubsection{Spray solution deposition at
downwind}\label{spray-solution-deposition-at-downwind}}

Data was fitted to a Bayesian linear mixed model using \emph{brm}
function. Spray solution deposition and distance were log-transformed to
meet linearity. A single model was fitted to each treatment. For each
model, treatments and distance were the fixed effects and block nested
within location random effects. Model family was gaussian and prior
distribution was set to student-t with mean 0.5, standard deviation 3
and 11 degrees of freedom. For clarification, intercepts, slopes were
back-transformed with \emph{exponential} function. Moreover, the linear
models fitted were used to predict the distance where no spray particle
deposition was detected (0 ηL cm\textsuperscript{-2}) for each
treatment, which was also back-transformed to m scale.

The area under the curve (AUC) was used to validate the linear models.
The spray solution deposition across distances within a experimental
unit were used to calculate the absolute AUC value. The AUC was
performed with \emph{audps} function from ``agricolae'' package. The AUC
is commonly used for plant disease progress\textsuperscript{5,6} but has
been used to calculate herbicide injury.\textsuperscript{7} Data was
fitted to a mixed model using \emph{brm} function. Treatments were the
fixed effects and block nested within location random effects. Model
family was gaussian and prior distribution was set to student-t with
mean 0.5, standard deviation 3 and 11 degrees of freedom. The posterior
summaries (mean and highest posterior density) were estimated with
\emph{emmeans} function from ``emmeans'' package. Treatment means were
compared using Bayes Factor (BF) as above-mentioned.

\hypertarget{dv-relative-span}{%
\subsubsection{Dv, Relative Span}\label{dv-relative-span}}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{spray-solution-deposition-at-inswath-and-upwind}{%
\subsection{Spray solution deposition at inswath and
upwind}\label{spray-solution-deposition-at-inswath-and-upwind}}

In general, Open sprayer treatments resulted in a more variable spray
particle deposition than Hood treatments (Figure 1). The inclusion of
either DRA or Water strongly impacted spray particle deposition inswath
for Open sprayer treatments, regardless nozzle type. The top and bottom
three treatments contained either DRA or Water, respectively. For
example, treatment DRA-Open-ULD resulted in the highest spray particle
deposition (1318.5 ηL cm\textsuperscript{-2}, Figure 1). In contrast,
911.2 nL cm\textsuperscript{-2} was the lowest spray solution
deposition, which was achieved with Water-Open-ULD treatment. Hood
sprayer treatments resulted in a more uniform spray particle deposition.
Furthermore, there were less than 0.29 ηL cm\textsuperscript{-2} spray
deposition at upwind with a strong evidence (BF \textless{} 0.25) of no
difference between all pairwise treatments contrasts (data not shown).

\hypertarget{spray-solution-deposition-at-downwind-1}{%
\subsection{Spray solution deposition at
downwind}\label{spray-solution-deposition-at-downwind-1}}

Treatments with highest intercepts, which is the amount of spray
particle deposition near the treated area, were Water-Open-AIXR (15.7 ηL
cm\textsuperscript{-2}), followed by DRA-Open-AIXR (15.7 ηL
cm\textsuperscript{-2}), and Water-Open-ULD (12.0 ηL
cm\textsuperscript{-2}; Figure 2). In contrast, DRA-Hood-TTI,
DRA-Hood-ULD and DRA-Hood-AIXR treatments resulted in the lowest
intercepts (\textless{} 2.0 ηL cm\textsuperscript{-2}). In addition,
there is evidence that treatments with Hood sprayer provided faster
decay of spray particle deposition (slopes; Figure 2). The treatments
with highest slope decay were Water-Hood-TTI (-0.50), DRA-Hood-TTI
(-0.48), DRA-Hood-ULD (-0.44), Water-Hood-AIXR (-0.43), Water-Hood-ULD
(-0.43), and DRA-Hood-AIXR (-0.39).

The predicted distance where no spray particle deposition was detected
varied upon treatments (Figure 3A). In general, Open sprayer treatments
resulted in spray particle deposition at longest distances. For example,
the distance of non-detectable spray particle deposition with Open
sprayer treatments varied from 9.9 m (DRA-Open-TTI) to 54.9 m with
DRA-Open-AIXR; whereas Hood sprayer treatments varied from 1.4 to 8.2 m
with DRA-Hood-TTI and Water-Hood-AIXR, respectively.

Similar trend was observed to AUC. Treatments with Hood or Open sprayer
strongly impacted on AUC values (Figure 3B). The highest AUC values were
Water-Open-AIXR (87.4), followed by Water-Open-ULD (72.9) and
DRA-Open-AIXR (72.6). In contrast, DRA-Hood-TTI (13.9), DRA-Hood-ULD
(14.3) and Water-Open-AIXR (21.9) resulted in lowest AUC values. The
impact of Open and Hood sprayer is demonstrated in treatments including
AIXR nozzle. There is a high difference in AUC (50.7) between
DRA-Open-AIXR vs DRA-Hood-AIXR (BF \textgreater{} 100). Moreover,
addition of DRA did reduced AUC values when comparing within fixed
factors, sprayer (Hood and Open) and nozzle (AIXR, TTI and ULD);
however, addition of DRA were not statistically different for some
contrasts, including DRA-Hood-TTI vs Water-Hood-TTI (BF = 0.91),
DRA-Open-TTI vs Water-Open-TTI (BF = 0.55), and DRA-Open-AIXR vs
Water-Open-AIXR (BF = 0.93).

\hypertarget{dv-relative-span-1}{%
\subsection{Dv, Relative Span}\label{dv-relative-span-1}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{spray-solution-deposition-at-inswath-and-upwind-1}{%
\subsection{Spray solution deposition at inswath and
upwind}\label{spray-solution-deposition-at-inswath-and-upwind-1}}

\hypertarget{spray-solution-deposition-at-downwind-2}{%
\subsection{Spray solution deposition at
downwind}\label{spray-solution-deposition-at-downwind-2}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

\hypertarget{conflict-of-interest-declaration}{%
\section{Conflict of Interest
Declaration}\label{conflict-of-interest-declaration}}

\hypertarget{tables-each-table-complete-with-title-and-footnotes}{%
\section{Tables (each table complete with title and
footnotes)}\label{tables-each-table-complete-with-title-and-footnotes}}

\hypertarget{figure-legends}{%
\section{Figure Legends}\label{figure-legends}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-LanguageEnvironmentStatistical2021}{}%
\CSLLeftMargin{1 }
\CSLRightInline{R: {A Language} and {Environment} for {Statistical
Computing}, {R Foundation for Statistical Computing}, {Vienna, Austria}
(2021).}

\leavevmode\hypertarget{ref-burknerBrmsPackageBayesian2017}{}%
\CSLLeftMargin{2 }
\CSLRightInline{Bürkner P-C, \textbf{Brms} : {An} {\emph{R}} {Package}
for {Bayesian Multilevel Models Using} {\emph{Stan}}, \emph{J Stat Soft}
\textbf{80} (2017).}

\leavevmode\hypertarget{ref-leeBayesianCognitiveModeling2014}{}%
\CSLLeftMargin{3 }
\CSLRightInline{Lee MD and Wagenmakers E-J, Bayesian {Cognitive
Modeling}: {A Practical Course}, {Cambridge University Press} (2014).}

\leavevmode\hypertarget{ref-kassBayesFactors1995}{}%
\CSLLeftMargin{4 }
\CSLRightInline{Kass RE and Raftery AE, Bayes {Factors}, \emph{Journal
of the American Statistical Association} \textbf{90}:773--795,
{{[}American Statistical Association, Taylor \& Francis, Ltd.{]}}
(1995).}

\leavevmode\hypertarget{ref-meenaAreaDiseaseProgress2011}{}%
\CSLLeftMargin{5 }
\CSLRightInline{Meena PD, Chattopadhyay C, Meena SS, and Kumar A, Area
under disease progress curve and apparent infection rate of {Alternaria}
blight disease of {Indian} mustard ({Brassica} juncea) at different
plant age, \emph{Archives of Phytopathology and Plant Protection}
\textbf{44}:684--693, {Taylor \& Francis} (2011).}

\leavevmode\hypertarget{ref-simkoAreaDiseaseProgress2011}{}%
\CSLLeftMargin{6 }
\CSLRightInline{Simko I and Piepho H-P, The {Area Under} the {Disease
Progress Stairs}: {Calculation}, {Advantage}, and {Application},
\emph{Phytopathology®} \textbf{102}:381--389, {Scientific Societies}
(2011).}

\leavevmode\hypertarget{ref-striegelSpraySolutionPH2021}{}%
\CSLLeftMargin{7 }
\CSLRightInline{Striegel S, Oliveira MC, Arneson N, Conley SP,
Stoltenberg DE, and Werle R, Spray solution {pH} and soybean injury as
influenced by synthetic auxin formulation and spray additives,
\emph{Weed Technol} \textbf{35}:113--127 (2021).}

\end{CSLReferences}

\end{document}