### 2021-08-17 analysis/bayes_droplet.Rmd
kok

---
title: "bayes_droplet"
author: "Maxwel Coura Oliveira"
date: "6/11/2021"
output: html_document
---

```{r}
library(tidyverse)
library(emmeans)
library(brms)
library(bayestestR)
library(modelbased)
library(ggtext)
library(patchwork)
library(ggthemes)
```

```{r}
readxl::read_excel("droplet_size_data.xlsx") %>% 
  janitor::clean_names() %>% 
  mutate(solution = str_replace_all(solution, "Intact", "DRA")) %>% 
  group_by(solution, nozzle) %>% 
  mutate(nozzle = str_remove_all(nozzle, '[0-9]+')) %>% 
  mutate(x200 = x200/100) %>% 
  mutate(rep = row_number()) -> droplet
```

# dv10

```{r message = FALSE}
model13 <- brm(`dv10` ~ solution * nozzle,
                  family = hurdle_gamma(link = "log", 
                                        link_shape = "log", 
                                        link_hu = "logit"),
               warmup = 1000, iter = 2000, chains = 4,
               prior = set_prior("student_t(0.5, 1, 2)", class = "b"),
                  data = droplet, save_pars = save_pars(all = TRUE))
```

```{r}
#library(bayestestR)
#bayesfactor_models(model11, model12, model13, denominator = model10)
# There are extreme evidence for model13 (interaction)
```

```{r}
droplet %>% 
  group_by(nozzle, solution) %>% 
  summarise(mean = mean(`dv10`), median = median(`dv10`))
```

```{r}
rg <- ref_grid(model13)
em <- emmeans(rg, ~ nozzle * solution, type = "response")
summary(em, point.est = mean)
```

```{r}
emmeans1 <- emmeans(model13, ~ nozzle * solution, cont="pairwise", type = "response")
```

```{r}
options(scipen = 999)
estimate_contrasts(model13, test = "bf") %>% 
  as_tibble() -> bf_dv10
write_csv(x = bf_dv10, file = "bf_dv10.csv")
```

# DV 50

```{r}
model23 <- brm(`dv50` ~ solution * nozzle,
                  family = hurdle_gamma(link = "log", 
                                        link_shape = "log", 
                                        link_hu = "logit"),
               warmup = 1000, iter = 2000, chains = 4,
               prior = set_prior("student_t(0.5, 1, 2)", class = "b"),
                  data = droplet, save_pars = save_pars(all = TRUE))
```

```{r}
# model comparison
#bayesfactor_models(model22, model23, denominator = model21)
```

```{r}
rg <- ref_grid(model23)
em <- emmeans(rg, ~ nozzle * solution, type = "response")
summary(em, point.est = mean)
```

```{r}
emmeans2 <- emmeans(model23, ~ nozzle * solution, cont="pairwise", type = "response")
```

```{r}
estimate_contrasts(model23, test = "bf")
```

```{r}
options(scipen = 999)
estimate_contrasts(model23, test = "bf") %>% 
  as_tibble() -> bf_dv50
#write_csv(x = bf_dv50, file = "bf_dv50.csv")
```

# DV 90

```{r}
model33 <- brm(`dv90` ~ solution * nozzle,
                  family = hurdle_gamma(link = "log", 
                                        link_shape = "log", 
                                        link_hu = "logit"),
               warmup = 1000, iter = 2000, chains = 4,
               prior = set_prior("student_t(0.5, 1, 2)", class = "b"),
                  data = droplet, save_pars = save_pars(all = TRUE))
```

```{r}
# model comparison
#bayesfactor_models(model32, model33, denominator = model31)
```

```{r}
rg <- ref_grid(model33)
em <- emmeans(rg, ~ nozzle * solution, type = "response")
summary(em, point.est = mean)
```

```{r}
emmeans3 <- emmeans(model33, ~ nozzle * solution, cont="pairwise", type = "response")
```

```{r}
estimate_contrasts(model33, test = "bf")
```

```{r}
estimate_contrasts(model33, test = "bf") %>% 
  as_tibble() -> bf_dv90
write_csv(x = bf_dv90, file = "bf_dv90.csv")
```

# RS

```{r}
model43 <- brm(rs ~ solution * nozzle,
                  family = hurdle_gamma(link = "log", 
                                        link_shape = "log", 
                                        link_hu = "logit"),
               warmup = 1000, iter = 2000, chains = 4,
               prior = set_prior("student_t(0.1, 0.5, 2)", class = "b"),
                  data = droplet, save_pars = save_pars(all = TRUE))
```

```{r}
# model comparison
#bayesfactor_models(model32, model33, denominator = model31)
```

```{r}
rg <- ref_grid(model43)
em <- emmeans(rg, ~ nozzle * solution, type = "response")
summary(em, point.est = mean)
```

```{r}
emmeans4 <- emmeans(model43, ~ nozzle * solution, cont= "pairwise", type = "response")
```

```{r}
estimate_contrasts(model43, test = "bf")
```

# <200

# RS

```{r}
model53 <- brm(`x200` ~ solution * nozzle,
                  family = Beta(link = "logit", link_phi = "log"),
               warmup = 1000, iter = 2000, chains = 4,
               prior = set_prior("student_t(0.1, 0.5, 2)", class = "b"),
                  data = droplet, save_pars = save_pars(all = TRUE))
```

```{r}
# model comparison
#bayesfactor_models(model32, model33, denominator = model31)
```

```{r}
rg <- ref_grid(model53)
em <- emmeans(rg, ~ nozzle * solution, type = "response")
summary(em, point.est = mean)
```

```{r}
emmeans5 <- emmeans(model53, ~ nozzle * solution, cont= "pairwise", type = "response")
```

```{r}
emmeans5$emmeans %>% 
  as_tibble() %>% 
  unite("trt", c("solution", "nozzle"), sep = " - ") %>% 
  ggplot(aes(x = fct_reorder(trt, response), y = response*100)) +
  geom_point() +
  geom_linerange(aes(ymin = lower.HPD*100, ymax = upper.HPD*100)) +
  coord_flip() +
  theme_classic() +
  labs(title = "<200 droptlet")
```

```{r}
estimate_contrasts(model53, test = "bf")
```

```{r}
emmeans1$emmeans %>% as_tibble() %>% mutate(type = "DV10") -> em1
emmeans2$emmeans %>% as_tibble() %>% mutate(type = "DV50") -> em2
emmeans3$emmeans %>% as_tibble() %>% mutate(type = "DV90") -> em3
```
-> emm
```{r}
em1 %>% 
  bind_rows(em2) %>% 
  bind_rows(em3) %>% 
  unite("trt", c("solution", "nozzle"), sep = "-") %>% 
  mutate(type = fct_recode(type,
                          "D<sub>v10</sub>" = "DV10",
                          "D<sub>v50</sub>" = "DV50",
                          "D<sub>v90</sub>" = "DV90")) -> emm
```

```{r}
droplet %>% 
  pivot_longer(3:5, names_to = "type", values_to = "response") %>%
  mutate(type = str_to_upper(type)) %>% 
  unite("trt", c("solution", "nozzle"), sep = "-") %>% 
  dplyr::select(-rs, -x200, -rep) %>% 
  mutate(type = fct_recode(type,
                          "D<sub>v10</sub>" = "DV10",
                          "D<sub>v50</sub>" = "DV50",
                          "D<sub>v90</sub>" = "DV90")) -> droplet1
```

```{r}
library(ggthemes)
emm %>% 
  ggplot(aes(x = fct_reorder(trt, response), y = response, color = trt)) +
  geom_point(size = 2) +
  geom_linerange(aes(ymin = lower.HPD, ymax = upper.HPD), size = 1.3) +
  geom_jitter(data = droplet1, alpha = 0.1, size = 2) +
  facet_grid(~ type) +
  coord_flip() +
  labs(x = NULL, y = "Droplet size (µm)") +
  scale_color_viridis_d(option = "C") +
  theme_few() + 
  theme(strip.text = element_markdown(size = 12),
        legend.position = "none") -> dv
```

```{r}
emmeans4$emmeans %>% as_tibble() -> em4
emmeans5$emmeans %>% as_tibble() -> em5
```

```{r}
droplet %>% 
  unite("trt", c("solution", "nozzle"), sep = "-") %>% 
  dplyr::select(trt, rs) %>% 
  rename(response = rs) -> droplet2
```

```{r}
em4 %>% 
  unite("trt", c("solution", "nozzle"), sep = "-") %>% 
  ggplot(aes(x = fct_reorder(trt, response), y = response, color = trt)) +
  geom_point(size = 2) +
  geom_linerange(aes(ymin = lower.HPD, ymax = upper.HPD), size = 1.3) +
  geom_jitter(data = droplet2, alpha = 0.1, size = 2) +
  coord_flip() +
  labs(x = NULL, y = "RS") +
  scale_color_viridis_d(option = "C") +
  theme_few() + 
  theme(strip.text = element_markdown(),
        legend.position = "none") -> rs
```

```{r}
droplet %>% 
  unite("trt", c("solution", "nozzle"), sep = "-") %>% 
  dplyr::select(trt, x200) %>% 
  rename(response = x200) -> droplet3
```

```{r}
(
em5 %>% 
  unite("trt", c("solution", "nozzle"), sep = "-") %>% 
  ggplot(aes(x = fct_reorder(trt, response*100), y = response*100, color = trt)) +
  geom_point(size = 2) +
  geom_linerange(aes(ymin = lower.HPD*100, ymax = upper.HPD*100), size = 1.3) +
  geom_jitter(data = droplet3, alpha = 0.1, size = 2) +
  coord_flip() +
  scale_y_continuous(breaks = c(0, 2, 4, 6, 8)) +
  labs(x = NULL, y = "% drifable fines < 200 µm") +
  scale_color_viridis_d(option = "C") +
  theme_few() + 
  theme(strip.text = element_markdown(),
        legend.position = "none") -> x200
)
```

```{r}
rs + x200 -> fig
dv / fig -> fig1

ggsave("Figure 4.png", width = 9, height = 6)
```