### 2021-08-19 analysis/droplet.Rmd

---
title: "droptlet"
author: "Maxwel Coura Oliveira"
date: "6/10/2021"
output: html_document
---

```{r}
library(tidyverse)
library(glmmTMB)
library(emmeans)
library(betareg)
library(lme4)
library(multcomp)
library(lmerTest)
library(car)
library(ggthemes)
library(ggrepel)
library(ggtext)
library(patchwork)
```

```{r}
library(extrafont)
library(showtext)
showtext_auto()
font_add_google("Lato", ## name of Google font
                "lato")  ## name that will be used in R
font_add_google("Roboto", "roboto")
```

```{r}
#readxl::read_excel("droplet_data.xlsx")
```

```{r}
readxl::read_excel("droplet_size_data.xlsx") %>% 
  janitor::clean_names() %>% 
  mutate_if(is_character, as_factor) %>% 
  group_by(solution, nozzle) %>% 
  mutate(nozzle = fct_recode(nozzle, 
                             "ULD12003" = "ULD11003")) %>% 
  mutate(x200 = x200/100) %>% 
  mutate(rep = row_number()) %>% 
  mutate(solution = str_replace_all(solution, "Intact", "DRA"),
         solution = str_replace_all(solution, "Water", "")) %>% 
  group_by(solution, nozzle) %>% 
  mutate(nozzle = str_remove_all(nozzle, '[0-9]+')) %>% 
  unite("trt", c(solution, nozzle), sep = "-", remove = FALSE) %>% 
  mutate(trt = str_remove(trt, "^[^a-zA-Z]+")) -> droplet
```

# DV10

```{r}
model1 <- glm(`dv10` ~ solution * nozzle,
                  family = Gamma(link = "inverse"),
                  data = droplet)
```

```{r}
Anova(model1)
```

```{r}
emmeans1 <- emmeans(model1, ~ nozzle * solution, type = "response")
```

```{r}
cld1 <- cld(emmeans1, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>% 
  as_tibble() %>% 
  mutate(type = "Dv10")
cld1
```

# DV50

```{r}
model2 <- glm(`dv50` ~ solution * nozzle,
                  family = Gamma(link = "inverse"),
                  data = droplet)
```

```{r}
Anova(model2)
```

```{r}
emmeans2 <- emmeans(model2, ~ nozzle * solution, type = "response")
```

```{r}
cld2 <- cld(emmeans2, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>% 
  as_tibble() %>% 
  mutate(type = "Dv50")

cld2
```

# DV90

```{r}
model3 <- glm(`dv90` ~ solution * nozzle,
                  family = Gamma(link = "inverse"),
                  data = droplet)
```

```{r}
Anova(model3)
```

```{r}
emmeans3 <- emmeans(model3, ~ nozzle * solution, cont="pairwise", type = "response")
```

```{r}
cld3 <- cld(emmeans3$emmeans, alpha=0.05, Letters=letters, 
            adjust="none", reversed = TRUE) %>% 
  as_tibble() %>% 
  mutate(type = "Dv90")
cld3
```

```{r}
cld1 %>% 
  bind_rows(cld2) %>% 
  bind_rows(cld3) %>% 
  group_by(solution, nozzle) %>% 
  mutate(nozzle = str_remove_all(nozzle, '[0-9]+')) %>% 
  unite("trt", c(solution, nozzle), sep = "-") %>% 
  mutate(trt = str_remove(trt, "^[^a-zA-Z]+")) %>% 
  mutate(response = round(response, 1)) -> dv
```

```{r}
library(scales)
show_col(brewer_pal(palette = "Set1")(5))
show_col(wsj_pal()(9))
```

```{r}
dv
droplet %>% 
  mutate(solution = str_replace_all(solution, "Intact", "DRA")) %>%
  pivot_longer(cols = c("dv10", "dv50", "dv90"), names_to = "type",
                             values_to = "dv") %>% 
  mutate(type = str_to_title(type)) -> droplet1
```

```{r}
dv %>% 
  mutate(type = fct_recode(type, 
                           "D<sub>v10</sub>" = "Dv10",
                           "D<sub>v50</sub>" = "Dv50",
                           "D<sub>v90</sub>" = "Dv90")) %>% 
  ggplot(aes(x = fct_reorder(trt, response), 
             y = response, 
             color = trt, 
             shape = type)) +
  geom_point(size = 2, position = position_dodge2(width = 0.4)) +
  geom_linerange(aes(ymin = asymp.LCL, ymax = asymp.UCL),
                 position = position_dodge2(width = 0.4), 
                 size = 1.3) +
  geom_jitter(data = droplet1 %>% 
                mutate(type = fct_recode(type, 
                           "D<sub>v10</sub>" = "Dv10",
                           "D<sub>v50</sub>" = "Dv50",
                           "D<sub>v90</sub>" = "Dv90")) %>% 
                rename(response = dv), 
                position = position_dodge2(width = 0.4),
              mapping = aes(color = trt, shape = type), alpha = 0.1, size = 2) +
  coord_flip() +
  geom_text_repel(aes(label = response, 
                         x = trt, color = trt, shape = type),
                    size = 8,
                    force_pull = 0,
                    direction = "y",
                    point.padding = NA,
                    min.segment.length = 14,
                    vjust = 0.5,
                    position = position_dodge2(width = 0.4),
                    seed = 455) +
#  annotate("text", x = 6, y = 300, label = "Dv10", size = 10,
#           color = "#E41A1C", fontface = "bold") +
#  annotate("text", x = 5.5, y = 800, label = "Dv50", size = 10,
#           color = "#377eb8", fontface = "bold") +
#  annotate("text", x = 4.5, y = 1250, label = "Dv90", size = 10, 
#           color = "#4daf4a", fontface = "bold") +
#  annotate("text", x = 1.2, y = 820, 
#           label = "There are differences \namong all pairwise trt \ncomparisons within \neach Dv (P <.0001)", size = 7.5, hjust = 0, lineheight = 0.3,
#           color = 1, fontface = "italic") +
  labs(x = NULL, y = "Droplet size (µm)", color = NULL, shape = NULL) +
  scale_color_viridis_d(option = "H") +
  scale_y_continuous(limits = c(0, 1500), breaks = seq(0, 1500, 300)) +
  scale_shape_manual(values = c(8,9,10)) +
  theme_few() + 
  theme(legend.position = c(0.85, 0.15),
        legend.text = element_markdown(size = 20),
        legend.background = element_blank(),
        axis.title = element_markdown(size = 30),
        axis.text = element_markdown(size = 25)) +
  guides(color = "none") -> fig1
```

# RS

```{r}
model4 <- glm(rs ~ solution * nozzle,
                  family = Gamma(link = "inverse"),
                  data = droplet)
```

```{r}
Anova(model4)
```

```{r}
emmeans4 <- emmeans(model4, ~ nozzle * solution, cont = "pairwise", type = "response")
```

```{r}
cld4 <- cld(emmeans4$emmeans, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>% 
  as_tibble() %>% 
  mutate(solution = str_replace_all(solution, "Intact", "DRA")) %>% 
  group_by(solution, nozzle) %>% 
  mutate(nozzle = str_remove_all(nozzle, '[0-9]+')) %>% 
  unite("trt", c(solution, nozzle), sep = "-") %>% 
  mutate(trt = str_remove(trt, "^[^a-zA-Z]+")) %>%
  mutate(response = round(response, 2))
cld4
```

```{r}
cld4 %>% 
  ggplot(aes(x = fct_reorder(trt, response), y = response, color = trt)) +
  geom_point(size = 2, position = position_dodge2(width = 0.4)) +
  geom_linerange(aes(ymin = asymp.LCL, ymax = asymp.UCL),
                 position = position_dodge2(width = 0.4), size = 1.3) +
  coord_flip() +
  geom_jitter(data = droplet1 %>% 
                rename(response = rs), 
              mapping = aes(color = trt), alpha = 0.1, size = 2) +
  geom_text_repel(aes(label = response, 
                         x = trt),
                    size = 8,
                    force_pull = 0,
                    direction = "y",
                    point.padding = NA,
                    min.segment.length = 14,
                    vjust = 0.5,
                    position = position_dodge2(width = 0.4),
                    seed = 455) +
  annotate("text", x = 5, y = 0.1, label = "There are differences \namong all pairwise trt  \ncomparisons (P <.0001) \nin figures A, B and C", size = 9, hjust = 0, lineheight = 0.3,
           color = 1, fontface = "italic") +
  labs(x = NULL, y = "Relative span", color = NULL) +
  scale_color_viridis_d(option = "H") +
  scale_y_continuous(limits = c(0, 1.2), breaks = seq(0, 1.2, 0.2)) +
  theme_few() + 
  theme(legend.position = "none",
        axis.title = element_markdown(size = 30),
        axis.text = element_markdown(size = 25)) -> fig2
```

# x200

```{r}
model5 <- betareg(x200 ~ solution * nozzle, 
                  link = "logit",
                  data = droplet)
```

```{r}
Anova(model5)
```

```{r}
emmeans5 <- emmeans(model5, ~ nozzle * solution, type = "response")
```

```{r}
cld5 <- cld(emmeans5, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>% 
  as_tibble() %>% 
  mutate(solution = str_replace_all(solution, "Intact", "DRA")) %>% 
  group_by(solution, nozzle) %>% 
  mutate(nozzle = str_remove_all(nozzle, '[0-9]+')) %>% 
  unite("trt", c(solution, nozzle), sep = "-") %>% 
  mutate(trt = str_remove(trt, "^[^a-zA-Z]+")) %>%
  mutate(emmean = round(emmean,3))
cld5
```

```{r}
cld5 %>% 
  ggplot(aes(x = fct_reorder(trt, emmean*100), y = emmean*100, color = trt)) +
  geom_point(size = 2, position = position_dodge2(width = 0.4)) +
  geom_linerange(aes(ymin = asymp.LCL*100, ymax = asymp.UCL*100),
                 position = position_dodge2(width = 0.4), size = 1.3) +
  coord_flip() +
  geom_jitter(data = droplet1 %>% 
                rename(emmean = x200), 
              mapping = aes(color = trt), alpha = 0.1, size = 2) +
  geom_text_repel(aes(label = emmean*100, 
                         x = trt),
                    size = 8,
                    force_pull = 0,
                    direction = "y",
                    point.padding = NA,
                    min.segment.length = 14,
                    vjust = 0.5,
                    position = position_dodge2(width = 0.4),
                    seed = 455) +
#  annotate("text", x = 2, y = 3, label = "There are differences \namong all pairwise trt \ncomparisons (P <.0001)", size = 9, hjust = 0, lineheight = 0.3,
#           color = 1, fontface = "italic") +
  labs(x = NULL, y = "% driftable fines < 200 µm", color = NULL) +
  scale_color_viridis_d(option = "H") +
  scale_y_continuous(limits = c(0, 8), breaks = seq(0, 8, 2)) +
  theme_few() + 
  theme(legend.position = "none",
        axis.title = element_markdown(size = 30),
        axis.text = element_markdown(size = 25)) -> fig3
```

```{r}
fig1 + fig2 + fig3 +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 35, hjust = 0, vjust = 0))

ggsave("Figure 4.png", width = 11, height = 4)
```