### 2021-08-19 supp_file/table.Rmd

---
title: "Supplemental File"
author: "Maxwel Coura Oliveira"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)
```

```{r message=FALSE}
# Packages
options(scipen = 999)
library(tidyverse)
library(reactable)
library(webshot2)
library(reactablefmtr)
```

$$\\[0.25in]$$
This is a supplemental material for the manuscript entitled: "XXXX". The Bayes Factor (BF) interpretation of Table 1 is based on Lee and Wagenmakers (2013). The H0 and H1 mean that treatments are the same or different, respectively.

$$\\[0.1in]$$

```{r}
read_csv("data/bf.csv") %>% 
  reactable(., theme = fivethirtyeight(), 
            filterable = FALSE,
            resizable = TRUE,
            searchable = FALSE,
            wrap = TRUE,
            defaultPageSize = 11,
            defaultColDef = colDef(minWidth = 200)) %>% 
  add_title("Table 1. Bayes Factor interpretation", font_size = 15)
```

$$\\[0.1in]$$

```{r}
read_csv("data/bf_dv10.csv") %>% 
  mutate(Parameter = str_remove_all(Parameter, "\\(|\\)")) %>% 
  mutate(log_BF = as.double(log_BF)) %>% 
  rename(BF = log_BF) %>% 
  mutate(BF = exp(BF)) %>% 
  mutate(Parameter = str_remove_all(Parameter, "Water ")) %>% 
  mutate_if(is.double, ~ round(., 2)) %>% 
  reactable(., theme = fivethirtyeight(),
            filterable = FALSE,
            resizable = TRUE,
            pagination = FALSE,
            wrap = TRUE,
            searchable = FALSE,
            defaultPageSize = 20,
             showSortable = TRUE,
            columns = list(
            BF = colDef(
              #cell = #function(value) {
#                if (value > 1) paste0("+", value) 
#                else value
#                },
              style = function(value) {
                color <- if (value > 1) {
                  "#008000"
                  } else if (value < 1) {
                    "#e00000"
                  }
                list(fontWeight = 600, color = color)}),
#            BF = colDef(align = "center",
#                        cell = color_tiles(., colors = ifelse(. > 1.00, "blue",
#                                                         "red"))),
            Parameter = colDef(align = "center", filterable = TRUE,
                                cell = data_bars_pos_neg(., 
                                                         colors = c("#ff3030", "#1e90ff"))))) %>% 
  add_title("Table 2. Treatment contrasts for DV10. Bayes Factor is used to test contrast hypothesis", font_size = 15)
```

```{r}
read_csv("data/bf_dv10.csv") %>% 
  mutate(Parameter = str_remove_all(Parameter, "\\(|\\)")) %>% 
  mutate(log_BF = as.double(log_BF)) %>% 
  rename(BF = log_BF) %>% 
  mutate(BF = exp(BF)) %>% 
  mutate_if(is.double, ~ round(., 2)) %>% 
  mutate(Parameter = str_remove_all(Parameter, "Water ")) %>% 
  reactable(., theme = fivethirtyeight(),
            filterable = FALSE,
            resizable = TRUE,
            pagination = FALSE,
            wrap = TRUE,
            searchable = FALSE,
            defaultPageSize = 20,
             showSortable = TRUE,
            columns = list(
            BF = colDef(
              #cell = #function(value) {
#                if (value > 1) paste0("+", value) 
#                else value
#                },
              style = function(value) {
                color <- if (value > 1) {
                  "#008000"
                  } else if (value < 1) {
                    "#e00000"
                  }
                list(fontWeight = 600, color = color)}),
#            BF = colDef(align = "center",
#                        cell = color_tiles(., colors = ifelse(. > 1.00, "blue",
#                                                         "red"))),
            Parameter = colDef(align = "center", filterable = TRUE,
                                cell = data_bars_pos_neg(., 
                                                         colors = c("#ff3030", "#1e90ff"))))) %>% 
  add_title("Table 3. Treatment contrasts for DV50. Bayes Factor is used to test contrast hypothesis", font_size = 15)
```

```{r}
read_csv("data/bf_dv10.csv") %>% 
  mutate(Parameter = str_remove_all(Parameter, "\\(|\\)")) %>% 
  mutate(log_BF = as.double(log_BF)) %>% 
  rename(BF = log_BF) %>% 
  mutate(BF = exp(BF)) %>% 
  mutate_if(is.double, ~ round(., 2)) %>% 
  mutate(Parameter = str_remove_all(Parameter, "Water ")) %>% 
  reactable(., theme = fivethirtyeight(),
            filterable = FALSE,
            resizable = TRUE,
            pagination = FALSE,
            wrap = TRUE,
            searchable = FALSE,
            defaultPageSize = 20,
             showSortable = TRUE,
            columns = list(
            BF = colDef(
              #cell = #function(value) {
#                if (value > 1) paste0("+", value) 
#                else value
#                },
              style = function(value) {
                color <- if (value > 1) {
                  "#008000"
                  } else if (value < 1) {
                    "#e00000"
                  }
                list(fontWeight = 600, color = color)}),
#            BF = colDef(align = "center",
#                        cell = color_tiles(., colors = ifelse(. > 1.00, "blue",
#                                                         "red"))),
            Parameter = colDef(align = "center", filterable = TRUE,
                                cell = data_bars_pos_neg(., 
                                                         colors = c("#ff3030", "#1e90ff"))))) %>% 
  add_title("Table 4. Treatment contrasts for DV90. Bayes Factor is used to test contrast hypothesis", font_size = 15)
```

```{r}
read_csv("data/bf_rs.csv") %>% 
  mutate(Parameter = str_remove_all(Parameter, "\\(|\\)")) %>% 
  mutate(log_BF = as.double(log_BF)) %>% 
  rename(BF = log_BF) %>% 
  mutate(BF = exp(BF)) %>% 
  mutate_if(is.double, ~ round(., 2)) %>% 
  mutate(Parameter = str_remove_all(Parameter, "Water ")) %>% 
  reactable(., theme = fivethirtyeight(),
            filterable = FALSE,
            resizable = TRUE,
            pagination = FALSE,
            wrap = TRUE,
            searchable = FALSE,
            defaultPageSize = 20,
             showSortable = TRUE,
            columns = list(
            BF = colDef(
              #cell = #function(value) {
#                if (value > 1) paste0("+", value) 
#                else value
#                },
              style = function(value) {
                color <- if (value > 1) {
                  "#008000"
                  } else if (value < 1) {
                    "#e00000"
                  }
                list(fontWeight = 600, color = color)}),
#            BF = colDef(align = "center",
#                        cell = color_tiles(., colors = ifelse(. > 1.00, "blue",
#                                                         "red"))),
            Parameter = colDef(align = "center", filterable = TRUE,
                                cell = data_bars_pos_neg(., 
                                                         colors = c("#ff3030", "#1e90ff"))))) %>% 
  add_title("Table 5. Treatment contrasts for relative span. Bayes Factor is used to test contrast hypothesis", font_size = 15)
```

```{r}
read_csv("data/bf_duz.csv") %>% 
  mutate(Parameter = str_remove_all(Parameter, "\\(|\\)")) %>% 
  mutate(log_BF = as.double(log_BF)) %>% 
  rename(BF = log_BF) %>% 
  mutate(BF = exp(BF)) %>% 
  mutate_if(is.double, ~ round(., 2)) %>% 
  mutate(Parameter = str_remove_all(Parameter, "Water ")) %>% 
  reactable(., theme = fivethirtyeight(),
            filterable = FALSE,
            resizable = TRUE,
            pagination = FALSE,
            wrap = TRUE,
            searchable = FALSE,
            defaultPageSize = 20,
             showSortable = TRUE,
            columns = list(
            BF = colDef(
              #cell = #function(value) {
#                if (value > 1) paste0("+", value) 
#                else value
#                },
              style = function(value) {
                color <- if (value > 1) {
                  "#008000"
                  } else if (value < 1) {
                    "#e00000"
                  }
                list(fontWeight = 600, color = color)}),
#            BF = colDef(align = "center",
#                        cell = color_tiles(., colors = ifelse(. > 1.00, "blue",
#                                                         "red"))),
            Parameter = colDef(align = "center", filterable = TRUE,
                                cell = data_bars_pos_neg(., 
                                                         colors = c("#ff3030", "#1e90ff"))))) %>% 
  add_title("Table 6. Treatment contrasts for % driftable fines < 200. Bayes Factor is used to test contrast hypothesis", font_size = 15)
```

The treatment contrasts of inswath spray solution deposition are compared (Table 2). See the BF and make an interpretation based on Table 1.

$$\\[0.1in]$$

```{r}
read_csv("data/inswath_bf.csv") %>%
  mutate(Parameter = str_remove_all(Parameter, "\\(|\\)")) %>% 
  mutate(log_BF = as.double(log_BF)) %>% 
  rename(BF = log_BF) %>% 
  mutate(BF = exp(BF)) %>% 
  mutate_if(is.double, ~ round(., 2)) %>% 
  reactable(., theme = fivethirtyeight(),
            filterable = FALSE,
            resizable = TRUE,
            pagination = FALSE,
            wrap = TRUE,
#            groupBy = "Level1",
            searchable = FALSE,
            defaultPageSize = 20,
            showSortable = TRUE,
            columns = list(
            BF = colDef(
              #cell = #function(value) {
#                if (value > 1) paste0("+", value) 
#                else value
#                },
              style = function(value) {
                color <- if (value > 1) {
                  "#008000"
                  } else if (value < 1) {
                    "#e00000"
                  }
                list(fontWeight = 600, color = color)}),
            Parameter = colDef(align = "center", filterable = TRUE, 
                                cell = data_bars_pos_neg(., 
                                                         colors = c("#ff3030", "#1e90ff"))))) %>% 
  add_title("Table 7. Treatment contrasts for in-swath spray solution deposition. Bayes Factor is used to test contrast hypothesis", font_size = 15)
```

$$\\[0.1in]$$
The treatment contrasts of Area Under the Curve are compared (Table 3). See the BF and make an interpretation based on Table 1.
$$\\[0.1in]$$

```{r}
read_csv("data/auc_bf.csv") %>% 
  mutate(Parameter = str_remove_all(Parameter, "\\(|\\)")) %>% 
  mutate(log_BF = as.double(log_BF)) %>% 
  mutate_if(is.double, ~ round(., 2)) %>% 
  rename(BF = log_BF) %>% 
  mutate(BF = exp(BF)) %>% 
  reactable(., theme = fivethirtyeight(),
            filterable = FALSE,
            resizable = TRUE,
            pagination = FALSE,
            wrap = TRUE,
            searchable = FALSE,
            defaultPageSize = 20,
             showSortable = TRUE,
            columns = list(
            BF = colDef(
              #cell = #function(value) {
#                if (value > 1) paste0("+", value) 
#                else value
#                },
              style = function(value) {
                color <- if (value > 1) {
                  "#008000"
                  } else if (value < 1) {
                    "#e00000"
                  }
                list(fontWeight = 600, color = color)}),
#            BF = colDef(align = "center",
#                        cell = color_tiles(., colors = ifelse(. > 1.00, "blue",
#                                                         "red"))),
            Parameter = colDef(align = "center", filterable = TRUE,
                                cell = data_bars_pos_neg(., 
                                                         colors = c("#ff3030", "#1e90ff"))))) %>% 
  add_title("Table 8. Treatment contrasts for Area Under the Curve. Bayes Factor is used to test contrast hypothesis", font_size = 15)
```

#### Reference

- Lee M.D. and Wagenmakers E.J. (2014) Bayesian Cognitive Modeling. Cambridge University Press. ISBN: 9781107018457, 1107018455