### 2021-08-31 analysis/combined.Rmd
nbn

---
title: "combined"
author: "Maxwel Coura Oliveira"
date: "5/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	fig.align = "center",
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	include = FALSE
)
```

```{r message = FALSE}
# Packages
options(scipen = 999)
library(rstanarm)
library(brms)
library(scales)
library(tidyverse)
library(emmeans)
library(ggthemes)
library(ggpmisc)
library(patchwork)
library(ggtext)
library(ggeffects)
library(agricolae)
library(extrafont)
library(showtext)
showtext_auto()
font_add_google("Lato", ## name of Google font
                "lato")  ## name that will be used in R
#font_add_google("Roboto", "roboto")
```

```{r}
new_dt <- read_csv("deposition.csv")
```

```{r}
new_dt %>% 
  mutate(solution= str_replace(solution, "Water", "")) %>% 
  unite("trt", c("solution", "sprayer", "nozzle"), 
        sep = "-", remove = FALSE) %>% 
  mutate(trt = str_replace(trt, "Intact", "DRA")) %>% 
  mutate(trt = str_remove(trt, "^[^a-zA-Z]+")) -> new_dt  
```

## Deposition inswath

```{r}
### Filter deposition dataset inswath
inswath <-  new_dt %>% 
  filter(type == "inswath") %>% 
  filter(!is.na(deposition))
```

```{r}
### Test ANOVA assumptions
#### Normal distribution
inswath %>% 
  ggplot(aes(x = deposition)) +
  geom_histogram()
# data has normal distribution
```

```{r}
#### Homogeneity of variances
bartlett.test(deposition ~ location, data = inswath)
# can't compare locations
```

```{r eval=FALSE}
### Data analysis
#### Model

#model1 <- lmer(deposition ~  trt +
#                (1|block/location), 
#              data = inswath)
```

```{r eval=FALSE}
#isSingular(model1)
```

```{r}
brm(deposition ~  trt +
                (1|block/location), 
    data = inswath,
    control = list(adapt_delta = 0.95),
    family = gaussian(),
    warmup = 1000, iter = 2000, chains = 4,
    prior = set_prior("student_t(0.5, 3, 11)", class = "b")) -> model1
#    prior = set_prior("normal(0, 100)", class = "b")) -> model1
```

```{r}
summary(model1, waic = TRUE)
```

```{r}
#plot(model1)
```

```{r}
rg <- ref_grid(model1)
em <- emmeans(rg, "trt")
summary(em, point.est = mean)

epred <- emmeans(model1, "trt", dpar = "mean", cont="pairwise")
summary(epred$emmeans, point.est = mean)
```

```{r}
#https://easystats.github.io/bayestestR/articles/bayes_factors.html
#https://www.statisticshowto.com/bayes-factor-definition/
groups <- emmeans(model1, ~ trt)
group_diff <- pairs(groups)
(groups_all <- rbind(groups, group_diff))

# pass the original model via prior

library(modelbased)
estimate_contrasts(model1, test = "bf")
```

```{r}
#https://easystats.github.io/blog/posts/bayestestr_emmeans/
inswath_main <- pairs(emmeans(model1, ~ trt))

describe_posterior(inswath_main,
                   centrality = "mean", dispersion = TRUE,
                   ci = .95, ci_method = "hdi",
                   test = c("bayesfactor"),
                   bf_prior = model1) %>% 
  tibble() -> inswath_bf
readr::write_csv(path = "../supp_file/data/inswath_bf.csv", x = inswath_bf)
```

```{r}
#estimate_contrasts(model1, test = "bf") -> inswath_bf
#inswath_bf %>% 
#  tibble() -> inswath_bf
#readr::write_csv(path = "../supp_file/data/inswath_bf.csv", x = inswath_bf)
```

```{r}
library(ggrepel)
em %>% as_tibble() %>% 
  mutate(sprayer = if_else(str_detect(trt, "Open"), "Open", "Hood")) %>% 
  mutate(emmean = round(emmean, 1)) %>% 
  ggplot(aes(x = fct_reorder(trt, emmean), y = emmean,
             color = trt, shape = sprayer, label = emmean)) +
  geom_point(size = 3) +
  geom_jitter(data = inswath %>% 
                rename(emmean = deposition), alpha = 0.1) +
  scale_color_viridis_d(option = "H") +
  coord_flip() +
  geom_text_repel(  size = 8,
                    force_pull = 0,
                    direction = "y",
                    point.padding = NA,
                    min.segment.length = 14,
                    vjust = 0.5,
                    position = position_dodge2(width = 0.4),
                    seed = 455) +
  scale_shape_manual(values = c(16, 1)) +
  geom_linerange(aes(ymin = lower.HPD, ymax = upper.HPD), size = 1.2) +
  ylim(0, 2000) +
  theme_few() +
  labs(x = NULL, 
       y = expression(paste("In-swath spray deposition  (\u03B7L/cm"^"2",")"))
       ) +
  theme(legend.position = "none",
        axis.title = element_text(size = 30),
        axis.text = element_markdown(size = 25)) -> figA

  ggsave("Figure 1.png", width = 5, height = 4)
```

## Deposition upwind

```{r}
### Filter deposition dataset upwind
upwind <- new_dt %>% 
  filter(type == "upwind") %>% 
  filter(!is.na(deposition))
```

```{r}
### Test ANOVA assumptions
#### Normal distribution
upwind %>% 
  ggplot(aes(x = deposition)) +
  geom_histogram() 
# not normalized
```

```{r}
##### Log scale
upwind %>% 
  ggplot(aes(x = log(deposition))) +
  geom_histogram()
# yes!
```

```{r}
brm(deposition ~  trt +
                (1|block/location), 
    data = upwind,
    control = list(adapt_delta = 0.99),
    family = gaussian(),
    warmup = 1000, iter = 2000, chains = 4,
    prior = set_prior("student_t(0.5, 3, 11)", class = "b")) -> model2
```

```{r}
rg2 <- ref_grid(model2)
em2 <- emmeans(rg2, "trt", point.est = mean)
summary(em2, point.est = mean)

epred2 <- emmeans(model2, "trt", dpar = "mean", cont="pairwise", type = "response")
summary(epred2$emmeans, point.est = mean)
```

```{r}
#library(bayestestR)
#bayesfactor_parameters(groups_all, prior = model1)
```

```{r}
#https://easystats.github.io/bayestestR/articles/bayes_factors.html
#https://www.statisticshowto.com/bayes-factor-definition/
groups <- emmeans(model1, ~trt)
group_diff <- pairs(groups)
(groups_all <- rbind(groups, group_diff))

# pass the original model via prior

library(modelbased)
estimate_contrasts(model2, test = "bf") -> upwind_bf
upwind_bf %>% 
  tibble() -> upwind_bf
readr::write_csv(path = "../supp_file/data/upwind_bf.csv", x = upwind_bf)
```

```{r eval = FALSE}
em2 %>% as_tibble() %>% 
  ggplot(aes(x = fct_reorder(trt, emmean), y = emmean,
             color = trt)) +
  geom_point(size = 3) +
  geom_jitter(data = upwind %>% 
                rename(emmean = deposition), alpha = 0.05) +
  scale_color_viridis_d(option = "H") +
  coord_flip() +
  geom_linerange(aes(ymin = lower.HPD, ymax = upper.HPD), size = 1.2) +
#  ylim(0, 0.5) +
  theme_few() +
  labs(x = NULL, y = expression(paste("Spray particle deposition (\u03b7L/cm"^"2",")"))) +
  theme(legend.position = "none") -> figB

#  ggsave("Figure 1B.png", width = 10, height = 5)
```

```{r eval = FALSE}
library(patchwork)

figB + figA + plot_annotation(tag_levels = 'A') -> figC

ggsave("Figure 1.png", width = 10, height = 5)
```

# Deposition downwind

```{r}
## Filter data
data_dw <- new_dt %>% 
  filter(type == "downwind") %>% 
  mutate(pass = factor(pass)) %>% 
  group_by(location, pass, block) 
```

```{r}
## Visualize
data_dw %>% 
  ggplot(aes(x = distance_m, y = deposition, color = nozzle)) +
  geom_point() +
  facet_grid(~ location)
# some trt have rapid decrease and some trt have linear trend
```

```{r}
## Visualize (log scale)
data_dw %>% 
#  unite("trt", c("solution", "sprayer", "nozzle")) %>% 
  ggplot(aes(x = log(distance_m), y = log(deposition), color = nozzle)) +
  geom_point() +
  facet_grid(sprayer ~ solution) +
  geom_smooth(method = "lm") 
# better go with log scale
```

```{r}
## Normal distribution
data_dw %>% 
  ggplot(aes(x = log(deposition))) +
  geom_histogram()
```

```{r}
# Linear Model
## Arrange dataset
data_dw1 <- data_dw %>% 
  group_by(trt) %>% 
  nest()

data_dw1
```

```{r}
pr <- prior(normal(0, 1), class = 'b')
### Model function
model3 <- function(df){
  brm(log(deposition) ~ log(distance_m) + 
         (1|block/location), data = df,

       prior = pr, cores=4)
}
```

```{r}
### Model function
model3 <- function(df){
  brm(log(deposition) ~ log(distance_m) + 
         (1|block/location), data = df,
      control = list(adapt_delta = 0.95),
      family = gaussian(),
      warmup = 1000, iter = 2000, chains = 4,
    prior = set_prior("student_t(0.5, 3, 11)", class = "b"))
}
```

```{r warning=FALSE, message = FALSE, warning = FALSE}
### Fit model and confidence intervals
data_dw2 <- data_dw1 %>% 
  mutate(model = map(data, model3),
         summary = map(model, broom.mixed::tidy))
```

```{r}
### Get parameters from models
params <- data_dw2 %>% 
  unnest(summary) %>% 
  filter(effect == "fixed") %>% 
  mutate(estimate_exp = exp(estimate),
         estimate_exp = case_when(
           term == "log(distance_m)" ~ -estimate_exp,
           TRUE                      ~ estimate_exp
         )) %>% 
  dplyr::select(trt, term, estimate, std.error, conf.low, conf.high) %>% 
  mutate(term = fct_recode(term,
                           slope = "logdistance_m"),
         term = fct_recode(term,
                           Intercept = "(Intercept)")) %>% 
  mutate(estimate = exp(estimate),
         conf.low = exp(conf.low),
         conf.high = exp(conf.high))

params
```

```{r}
params %>%

  pivot_wider(id_cols = trt,
              names_from = "term", 
              values_from = "estimate") %>% 
  mutate(slope = - slope) -> distance_final
```

```{r}
distance_function <- function(int, slope, y = 0) {
  distance = (y - int) / slope
  return(distance)
}

distance_final1 <- distance_final %>% 
  mutate(distance_m = map2_dbl(Intercept, slope, distance_function),
         distance_m = round(distance_m, 10)) %>% 
  mutate( y = Intercept + slope * distance_m)
```

```{r}
linear <- function(a, b, d) {
  y = a + b*d
  return(y)
}

linear(distance_final1$Intercept, distance_final1$slope, distance_final1$distance_m)
```

### Intercept

```{r include=TRUE, eval = FALSE}
a <- params %>% 
  filter(term == "Intercept") %>% 
  ggplot(aes(x = fct_reorder(trt, estimate), y = estimate, color = trt)) +
  geom_point(size = 2, stroke = 1) +
  scale_color_viridis_d(option = "H") +
  geom_linerange(aes(ymin = conf.low, ymax = conf.high), size = 1.5) +
  coord_flip() +
  labs(y = expression(paste("Spray particle deposition (\u03b7L/cm"^"2",")")),
       x = NULL) +
  theme_few() +
  theme(legend.position = "none")
a
```

### Slope

```{r include=TRUE, eval = FALSE}}
b <- params %>% 
  filter(term == "slope") %>% 
  ggplot(aes(x = fct_reorder(trt, -estimate), 
             y = -estimate, color = trt)) +
  geom_point(size = 2, stroke = 1) +
  scale_color_viridis_d(option = "H") +
  geom_linerange(aes(ymin = -conf.low, ymax = -conf.high), size = 1.5) +
  coord_flip() +
  labs(y = expression(paste("Spray particle deposition (\u03b7L/cm"^"2",")")),
       x = NULL) +
  theme_few() +
  theme(legend.position = "none") 
b
```

```{r eval = FALSE}
a + b + plot_annotation(tag_levels = 'A') -> figZ

ggsave("Figure 2.png", width = 10, height = 5)
```

### Predicted downwind distance

```{r include=TRUE}
c <- distance_final1 %>% 
  mutate(trt = str_replace(trt, "Intact", "DRA")) %>% 
  mutate(hood = if_else(str_detect(trt, "Open"), "Open", "Hood")) %>% 
  mutate(distance_m = round(distance_m, 1)) %>% 
  ggplot(aes(x = fct_reorder(trt, distance_m), 
             y = distance_m,
             label = distance_m, color = trt, shape = hood)) +
  geom_point(size = 3, stroke = 1) +
  geom_text_repel(
                    size = 8,
                    force_pull = 0,
                    direction = "y",
                    point.padding = NA,
                    min.segment.length = 14,
                    vjust = 0.5,
                    position = position_dodge2(width = 0.4),
                    seed = 545) +
  scale_shape_manual(values = c(16, 1)) +
  geom_segment(aes(x= trt, xend= trt, y=0, yend=distance_m), size = 1) +
  coord_flip() +
  scale_color_viridis_d(option = "H") + 
  labs(y = "Predicted distance (m)", x = NULL) +
  ylim(0,60) +
  theme_few(base_family = "lato") +
  theme(legend.position = "none",
        legend.text = element_markdown(size = 20),
        legend.background = element_blank(),
        axis.title = element_markdown(size = 30),
        axis.text = element_markdown(size = 25)) 
c
```

```{r}
fig <- data_dw2 %>% 
  dplyr::select(trt, model)
```

```{r warning=FALSE, message = FALSE}
fig_model <- function(model){
  ggpredict(model, terms = "distance_m", back.transform = TRUE) %>% 
    as_tibble()
}

fig1 <- fig %>% 
  mutate(estimates = map(model, fig_model)) %>% 
  unnest(estimates) %>% 
  rename(distance_m = x) %>% 
  mutate_if(is_character, as_factor) 
```

```{r}
fig1 %>% 
  distinct(trt) %>%
  ungroup() %>% 
  mutate(id = row_number()) -> id_col

fig1 %>% 
  left_join(id_col) -> fig1
```

```{r eval=FALSE, include=FALSE}
library(sjstats)
performance::r2(fig1$model[[2]]) -> m1
m1
```

```{r eval=FALSE, include=FALSE}
rsquared <- function(trt){
  performance::r2(fig1$model[[trt]])$R2_Bayes
}
```

```{r eval=FALSE, include=FALSE}
fig1 %>% 
  mutate(rsquared = map(id, rsquared)) %>% 
  unnest(rsquared) -> fig1
```

```{r}
label <- distance_final %>% 
  mutate(Intercept = round(Intercept, 1),
         slope = round(slope, 2)) %>% 
  mutate(y = "y = ",
         x = "x") %>% 
  unite("formula", c("y", "Intercept", "slope"), sep = " ") %>% 
  unite("formula", c("formula", "x"), sep = "")
```

```{r}
fig2 <- fig1 %>% 
  left_join(label)
```

### Back-transformed linear models

```{r}
loc <- tibble(trt = "DRA-Open-TTI")
```

```{r include=TRUE}
fig2 %>% 
  mutate(trt = str_replace(trt, "Intact", "DRA")) %>% 
  mutate(hood = if_else(str_detect(trt, "Open"), "Open", "Hood")) %>% 
  ggplot(aes(x = distance_m, y = predicted, color = trt, 
             fill = trt, label = formula, shape = hood)) +
  scale_color_viridis_d(option = "H") + 
  scale_fill_viridis_d(option = "H") + 
  geom_line(fill = NA) +
  geom_point(alpha = 0.5) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
              alpha = 0.3, color = NA) +
  facet_wrap(~ trt, ncol = 3) +
  scale_shape_manual(values = c(16, 1)) + 
  theme_few() +
#  geom_text(data = loc, mapping = aes(x = 10, y = 20), 
#            label = "Intercept", inherit.aes = FALSE, size = 10) +
#  geom_text(data = loc, mapping = aes(x = 55, y = 10), 
##            label = "Slope", inherit.aes = FALSE, size = 10) +
  geom_label(aes(label = formula), x = 30, y = 15, color = "white", size = 12) +
#  geom_curve(data = loc, mapping = aes(x = 25, y = 17,
#             xend = 20, yend = 20),
#             arrow = arrow(length = unit(0.07, "inch")), 
#             color = 1,
#             size = 0.4, curvature = 0.3, inherit.aes = FALSE) +
#  geom_curve(data = loc, mapping = aes(x = 42, y = 13,
#             xend = 47, yend = 10),
#             arrow = arrow(length = unit(0.07, "inch")), 
#             color = 1,
#             size = 0.4, curvature = 0.3, inherit.aes = FALSE) +
  labs(x = "Distance (m)", 
       y = expression(paste("Spray particle deposition (\u03b7L/cm"^"2",")"))) +
  theme(legend.position = "none",
        legend.text = element_markdown(size = 20),
        legend.background = element_blank(),
        axis.title = element_markdown(size = 32),
        axis.text = element_markdown(size = 28),
        strip.background = element_blank(),
        strip.text = element_textbox(
          size = 32,
          face = "bold",
          color = "white", 
          fill = "#494949", 
          box.color = "#282728",
          halign = 0.5, linetype = 1, r = unit(5, "pt"), width = unit(1, "npc"),
          padding = margin(2, 0, 1, 0), 
          margin = margin(3, 3, 3, 3)))

ggsave("Figure 2.png", width = 6, height = 9)
```

### AUC

```{r}
(id_df <- data_dw %>% 
  distinct(pass, block, location) %>% 
  ungroup() %>% 
  mutate(id = as_factor(dplyr::row_number()))
)
```

```{r}
id_df1 <- data_dw %>% 
  left_join(id_df) 
```

```{r}
auc_cal <- function(number){
  df <- id_df1 %>% 
    filter(id == number)

  audps(df$deposition, df$distance_m, type="absolute") 
}

auc_data <- id_df1 %>% 
  group_by(id) %>% 
  nest() %>% 
  mutate(auc = map_dbl(id, auc_cal)) %>% 
  ungroup(id) %>% 
  unnest() %>% 
  filter(!is.na(auc)) %>% 
  mutate(trt = str_replace(trt, "Intact", "DRA")) 
```

```{r}
auc_data1 <- auc_data %>% 
  filter(distance_m == 1)
```

```{r}
auc_data1 %>% 
  ggplot(aes(x = log(auc))) + 
  geom_histogram()
```

```{r}
brm(auc ~  trt +
                (1|block/location), 
    data = auc_data1,
    control = list(adapt_delta = 0.99),
    family = "gaussian",
    warmup = 1000, iter = 2000, chains = 4,
   prior = set_prior("student_t(0.5, 3, 11)", class = "b")) -> auc_model
```

```{r}
rg3 <- ref_grid(auc_model)
em3 <- emmeans(rg3, "trt")
summary(em3, point.est = mean)
```

```{r}
plot(em3, ~ "trt", type="response")
```

```{r}
epred3 <- emmeans(auc_model, "trt", dpar = "mean", 
                  cont="pairwise")
summary(epred3$emmeans, point.est = mean, type = "response")
#https://easystats.github.io/bayestestR/articles/bayes_factors.html
```

```{r}
epred3$emmeans %>% as_tibble() %>% 
  mutate(sprayer = if_else(str_detect(trt, "Open"), "Open", "Hood")) %>% 
  mutate(emmean = round(emmean, 1)) %>% 
  ggplot(aes(x = fct_reorder(trt, emmean), y = emmean,
             color = trt, shape = sprayer, label = emmean)) +
  geom_point(size = 3) +
  geom_jitter(data = auc_data %>% 
                rename(emmean = auc), alpha = 0.02) +
  scale_color_viridis_d(option = "H") +
  geom_text_repel(
                    size = 8,
                    force_pull = 0,
                    direction = "y",
                    point.padding = NA,
                    min.segment.length = 14,
                    vjust = 0.5,
                    position = position_dodge2(width = 0.4),
                    seed = 545) +
  coord_flip() +
  scale_shape_manual(values = c(16, 1)) +
  geom_linerange(aes(ymin = lower.HPD, ymax = upper.HPD), size = 1.2) +
  scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 250)) +
  theme_few() +
  labs(x = NULL, y = "Area under the curve") +
  theme(legend.position = "none",
        legend.text = element_markdown(size = 20),
        legend.background = element_blank(),
        axis.title = element_markdown(size = 30),
        axis.text = element_markdown(size = 25)) -> figE
#figE

ggsave("Figure 3.png", width = 5, height = 4)
```

```{r}
#https://easystats.github.io/blog/posts/bayestestr_emmeans/
auc_main <- pairs(emmeans(auc_model, ~ trt))

describe_posterior(auc_main,
                   centrality = "mean", dispersion = TRUE,
                   ci = .95, ci_method = "hdi",
                   test = c("bayesfactor"),
                   bf_prior = auc_model) %>% 
  tibble() -> auc_bf
readr::write_csv(path = "../supp_file/data/auc_bf.csv", x = auc_bf)
```

```{r}
#estimate_contrasts(auc_model, test = "bf") -> auc_bf
#auc_bf %>% 
#  tibble() -> auc_bf
#readr::write_csv(path = "../supp_file/data/auc_bf.csv", x = auc_bf)
```

```{r eval = FALSE}
#library(modelbased)
#estimate_contrasts(auc_model, test = "bf") %>% 
#  kableExtra::kable()
```

```{r}
#c + figE +
#  plot_annotation(tag_levels = 'A') & 
#  theme(plot.tag = element_text(size = 35, hjust = 0, vjust = 0)) -> figU

#ggsave("Figure 3.png", width = 8, height = 4)
```